---
title: "Big Mac Index"
date: 2021-01-03
output: html_output
---

# TidyTuesday

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
library(ggrepel)
library(gganimate)
library(transformr)
library(widyr)
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2020-12-22")


big_mac <- tt$`big-mac` %>%
  rename(country = name)%>%
add_count(country, name = "country_total")%>%
  arrange(country, date)
```

Big Mac Prices over time
```{r}
big_mac%>%
  filter(country_total == max(country_total))%>%
  
  mutate(country = fct_reorder(country, local_price, function(.) last(.) / first(.)))%>%
  ggplot(aes(date, local_price, color = country))+
  geom_line()+
  expand_limits(y = 0)+
  facet_wrap(~ country, scales = "free_y")+
  theme(legend.position = "none")+
  labs(x= "Time", y= "Price of Big Mac in local currency")
```

Big Mac Inflation
```{r}
big_mac%>%
  filter(country_total == max(country_total))%>%
  group_by(country) %>%
  summarize(big_mac_inflation = last(local_price) / first(local_price))%>%
  arrange(desc(big_mac_inflation))%>%
  mutate(country = fct_reorder(country, big_mac_inflation))%>%
  ggplot(aes(big_mac_inflation, country))+
  geom_col()+
  geom_text(aes(label= paste0(round(big_mac_inflation,1))), hjust= 0)+
  expand_limits(x = 130)+
  scale_x_log10(breaks = c(1, 3, 10 ,30, 100))+
  labs(x = "Price of Big Mac in 2020 / Price of Big Mac in 2000" , y = "")
```

```{r}
big_mac%>%
  filter(country_total == max(country_total))%>%
  group_by(date)%>%
  mutate(usd_price = local_price[iso_a3 == "USA"])%>%
  mutate(big_mac_ex = local_price/usd_price)%>%
  select(date, iso_a3, country, local_price, dollar_ex, usd_price, big_mac_ex, usd_raw, gdp_dollar, usd_adjusted)
  
```

```{r}
 big_mac%>%
  filter(country_total == max(country_total))%>%
  select(date, iso_a3, country, local_price, dollar_ex, usd_raw, gdp_dollar, usd_adjusted)%>%
  filter(!is.na(gdp_dollar))%>%
  ggplot(aes(date, usd_raw))+
  geom_line()+
  expand_limits(y = 0)+
  facet_wrap(~ country, scales = "free_y")
```

Recalculated Big Mac Index
```{r}
 big_mac%>%
  group_by(date)%>%
  mutate(usd_price = local_price[iso_a3 == "USA"],
         us_gdp = gdp_dollar[iso_a3 == "USA"])%>%
ungroup()%>%
  filter(country == "Argentina",
         !is.na(gdp_dollar))%>%
  mutate(price_from_usd = usd_price * dollar_ex)%>%
  ggplot(aes(date, local_price))+
  geom_line(aes(color = "Price in local currency"))+
  geom_line(aes(y = price_from_usd  , color = "Price from USD"))
```

Adjusted Big Mac Index
```{r}
 big_mac%>%
  filter(country_total == max(country_total),
         country != "United States")%>%
  select(date, iso_a3, country, local_price, dollar_ex, usd_raw, gdp_dollar, usd_adjusted)%>%
  filter(!is.na(gdp_dollar))%>%
  mutate(country = fct_reorder(country, usd_adjusted))%>%
  ggplot(aes(date, usd_adjusted))+
  geom_line()+
  geom_hline(color= "blue", lty= 2, yintercept = 0)+
  expand_limits(y = 0)+
  labs(y = "Adjusted Big Mac Index relative to USD", x= "")+
  facet_wrap(~ country)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```

Raw Big Mac Index
```{r}
 big_mac%>%
  filter(country_total == max(country_total),
         country != "United States")%>%
  filter(!is.na(gdp_dollar))%>%
  mutate(country = fct_reorder(country, usd_raw))%>%
  ggplot(aes(date, usd_raw))+
  geom_line()+
  geom_hline(color= "blue", lty= 2, yintercept = 0)+
  expand_limits(y = 0)+
  labs(y = "Raw Big Mac Index relative to USD", x= "")+
  facet_wrap(~ country)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```

Why is it adjusted? --> More than based on the wealth of a country. 
```{r}
 big_mac%>%
    filter(country_total == max(country_total))%>%
    filter(date == max(date))%>%
  ggplot(aes(gdp_dollar, usd_raw))+
      geom_point()+
      geom_text_repel(aes(label = country))+
      geom_smooth(method = "lm")+
      labs(x = "GDP per capita (dollars)",
           y = "Raw Big Mac Index relative to USD")
```

In Indonesia big max are cheaper then expected based on the currency conversion rate. However this is not really suprising because Indonesia's GDP per Capital is close to the trend line. 
It makes more sense to have a look at Countries that are poor, but the Big Mac price is fairly expensive or rich countries with a fairly cheap big mac price.

GDP Adjusted, the trend line is fairly flat so better Idea of overpriced undervalued.

```{r}
 big_mac%>%
    filter(country_total == max(country_total))%>%
    filter(date == max(date))%>%
  ggplot(aes(gdp_dollar, usd_adjusted))+
      geom_point()+
      geom_text_repel(aes(label = country))+
      geom_smooth(method = "lm")+
      labs(x = "GDP per capita (dollars)",
           y = "Adjusted Big Mac Index relative to USD")
```

Barplot you can see HongKong is undervalued Thailand overvalued. 
```{r}
 big_mac%>%
    filter(country_total == max(country_total))%>%
    filter(date == max(date))%>%
    mutate(country = fct_reorder(country, usd_adjusted))%>%
    ggplot(aes(usd_adjusted, country))+
    geom_col()+
    labs( x ="Big Mac Index relative to USD (GDP     adjusted)", y = "")
```


Animated Graph of GDP Adjusted of the over priced.
```{r}
 big_mac%>%
    filter(country_total == max(country_total))%>%
    filter(!is.na(gdp_dollar))%>%
  ggplot(aes(gdp_dollar, usd_adjusted))+
      geom_point()+
      geom_text_repel(aes(label = country))+
      geom_smooth(method = "lm")+
      transition_time(date)+
      labs(x = "GDP per capita (dollars)",
           y = "Adjusted Big Mac Index relative to USD",
           title = "{frame_time}")
```


Animated graph of Barplot
```{r}
 big_mac%>%
    filter(!is.na(gdp_dollar))%>%
    filter(country_total == max(country_total))%>%
    mutate(country = fct_reorder(country, usd_adjusted))%>%
    ggplot(aes(usd_adjusted, country))+
    geom_col()+
    transition_time(date)+
    labs( x ="Big Mac Index (GDP adjusted)", y = "",
          title = "{frame_time}")
```

What Big Mac Prices tend to move together?

```{r}
big_mac%>%
  pairwise_cor(country, date, local_price, sort = TRUE)%>%
  filter(item1 == "United States")
```


# Credits
Most of the Ideas and code are from David Robinson ([YT](https://www.youtube.com/user/safe4democracy), Github[https://github.com/dgrtwo]) I mainly do the analysis on my own inspired by his videos in order to practice and learn new code pieces. Furthermore I like adding my own thoughts, so I post them here on my website.

# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


